### 2025-05-07 paper/paper.tex

In order to run training, navigate to the TrafficSignalManagement file in a terminal instance and run the provided command. \texttt{\texttt{python traffic\_signal\_nn/main.py --config traffic\_signal\_nn/config/config\_dqn\_city.ini train}}. Common errors include not having SUMO variables added to the path, and not having pip modules installed. Each episode will run, taking about a minute to simulate an hour of traffic. The reward will be printed to the console to let the user observe the loss, to see when the model explores and when it converges. The code can take hours at a time to run, and a demo can be provided if necessary, via youtube.

In order to benchmark the baseline for SUMO, the scripts benchmark.py and get\_queue\_length.py were developed. benchmark simply gets the results of the runtime through the file data/output/tripinfo.xml. It contains information about the trip, such as average trip time, average waiting time, and average delay, but does not contain information on average queue length. Before running benchmarks, it is important to run the get queue length script, which obtains the average queue lengths seperate from what SUMO provides. One must run the benchmark scenario provided through simulation.sumocfg within SUMO-gui to get the correct output.

>>>>

In order to run training, navigate to the TrafficSignalManagement file in a terminal instance and run the provided command. \texttt{\texttt{python traffic\_signal\_nn/main.py --config traffic\_signal\_nn/config/config\_dqn\_city.ini train}}. Common errors include not having SUMO variables added to the path, and not having pip modules installed. Each episode will run, taking about a minute to simulate an hour of traffic. The reward will be printed to the console to let the user observe the loss, to see when the model explores and when it converges. Parameters can be modified for training within traffic\_signal\_nn/config/config\_dqn\_city.ini, including the neural network parameters and the selected roads. The code can take hours at a time to run, and a demo can be provided if necessary, via youtube.

In order to benchmark the baseline for SUMO, the scripts benchmark.py and get\_queue\_length.py were developed. benchmark simply gets the results of the runtime through the file data/output/tripinfo.xml. It contains information about the trip, such as average trip time, average waiting time, and average delay, but does not contain information on average queue length. Before running benchmarks, it is important to run the get queue length script, which obtains the average queue lengths seperate from what SUMO provides. One must run the benchmark scenario provided through simulation.sumocfg within SUMO-gui to get the correct output. In order to run the benchmarks, simply run \texttt{python benchmark.py} after running the simulation, along with \texttt{python .\textbackslash get\_queue\_length.py --config .\textbackslash data\textbackslash simulation.sumocfg --output .\textbackslash data\textbackslash output\textbackslash queue\_length.csv} to determine the queue lengths.

In order to run the benchmarks with the model, one must run the trained model first to get the trip information. This can be ran through this command \texttt{python -m traffic\_signal\_nn.eval.run\_model --config traffic\_signal\_nn/config/config\_dqn\_city.ini}, which runs the trained model on the same situation as the previous benchmark, for consistency as mentioned in cross-validation. Just like without the neural network, the queue length is not obtained for the benchmarks, so one will need to run \texttt{python .\textbackslash traffic\_signal\_nn\textbackslash eval\textbackslash get\_model\_queue\_length.py --config .\textbackslash traffic\_signal\_nn\textbackslash config\textbackslash config\_dqn\_city.ini --output .\textbackslash data\textbackslash output\textbackslash queue\_length.csv} in order the get the updated queue lengths. Once all of this has been done, one can simply run \texttt{python benchmark.py} to get the benchmark results for the trained model.

Setup of the project can be rather time consuming, as mentioned, a video can provided to the reviewers running the script completely setup if necessary.
